### 2015-03-31 doc/proposal/paper.tex
Add proposal template

\documentclass[letterpaper,twocolumn,10pt]{article}

\usepackage{style, verbatim, url}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}
\usepackage{cleveref}
\usepackage{breakurl}

\makeatletter
\newcommand{\crefnames}[3]{
    \@for\next:=#1\do{
        \expandafter\crefname\expandafter{\next}{#2}{#3}
    }
}
\makeatother
\crefnames{part,chapter,section}{\S}{\S\S}

\begin{document}

\title{\Large \bf A Transparent Peer-to-Peer CDN}

\author{
    {\rm Anish Athalye}\\
    aathalye@mit.edu
    \and
    {\rm Ankush Gupta}\\
    ankush@mit.edu
    \and
    {\rm Katie Siegel}\\
    ksiegel@mit.edu
}

\maketitle

\thispagestyle{empty}

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Design}

\section{Goals}

{\footnotesize \bibliographystyle{ieeetr}
\bibliography{paper}}

\end{document}