### 2015-05-07 doc/paper/paper.tex
Fix command spacing

\documentclass[letterpaper,twocolumn,10pt]{article}

\usepackage{style, verbatim, url, graphicx, xspace}
\usepackage[colorlinks=true, citecolor=blue, linkcolor=blue]{hyperref}
\usepackage{caption}
\usepackage{cleveref, breakurl}

\makeatletter
\newcommand{\crefnames}[3]{
    \@for\next:=#1\do{
        \expandafter\crefname\expandafter{\next}{#2}{#3}
    }
}
\makeatother
\crefnames{part,chapter,section}{\S}{\S\S}

\newcommand{\zbf}{Zauberfl\"{o}te\xspace}
\newcommand{\projtitle}{\zbf: A Transparent Peer-to-Peer CDN}
\newcommand{\inclfigure}[3]{
    \begin{figure}
        \begin{centering}
            \includegraphics[width=.8\linewidth]{#2}
            \caption{#3}
            \label{#1}
        \end{centering}
    \end{figure}
}

\begin{document}

\title{\Large \bf \projtitle}

\author{
    {\rm Anish Athalye}\\
    aathalye@mit.edu
    \and
    {\rm Ankush Gupta}\\
    ankush@mit.edu
    \and
    {\rm Katie Siegel}\\
    ksiegel@mit.edu
}

\maketitle
\thispagestyle{empty}

\begin{abstract}
% TODO
\end{abstract}

\section{Introduction}
In today's day and age, it's very common for personal blogs and other small websites to gain popularity on websites such as Reddit. Since these websites usually experience much smaller loads, the hardware devoted to hosting them is typically not powerful enough to cope with a large amount of traffic. These sudden, unexpected bursts of traffic often result in websites loading at extremely slow speeds (or not loading at all) and is colloquially known as the "Reddit Hug of Death".

Typical means for dealing with this situation involve utilizing Content Delivery Networks (CDNs), but these are usually out of the budgetary constraints for websites which only experience such loads on an occasional basis.

Currently, peer-to-peer content delivery networks do exist, yet are usually
closed-source. PeerCDN~\cite{peercdn} is one such example found in
industry; it offloads site hosting to users of a site, significantly lowering strain on content delivery from a central source. \zbf is an open-sourced system
that will allow developers to easily integrate a peer-to-peer CDN to expedite
content delivery to clients even when operating under heavy load.

\zbf's design is based on the BitTorrent protocol~\cite{cohen:bittorrent}, where
clients requesting content will attempt to receive chunks of data from other
clients, rather than retrieving the entire file from a central server.

\section{\zbf API}

\zbf supports peer-to-peer loading for CSS (via \texttt{link} elements), Javascript (via \texttt{script} elements), and images (via \texttt{img} elements). In order for an existing website to utilize \zbf, the developer simply needs to make two changes to the DOM elements.

The first change that must be made is renaming the \texttt{src} or \texttt{href} attribute to  \texttt{data-zf-fallback}. \zbf uses the \texttt{data-zf-fallback} attribute in order to (1) prevent the browser from automatically fetching the resources, and (2) provide a URL to be used in the event that the peer-to-peer network is unable to serve the content.

The second attribute that must be present on a DOM element that is fetched by \zbf is the \texttt{data-zf-hash} attribute. The \texttt{data-zf-hash} attribute is the SHA1 hash of that resource, which can be computed by the developer, and is utilized by \zbf in order to ensure integrity of data fetched via peer-to-peer interaction.

\section{Implementation}

The \zbf architecture consists of several loosely-connected components (see
figure~\ref{fig:components}). At the highest level, there is a download manager that
provides a reliable download service. It relies on the tracker interface and
the connection manager. The tracker interface facilitates communication with
the central tracker for information about peers and file sizes, and the
connection manager provides an interface for peer-to-peer message transfer.

Our components rely on two standard APIs provided by modern browsers,
WebSockets~\cite{w3c:websocket} for client-server communication and
WebRTC~\cite{w3c:webrtc} for peer-to-peer communication.

\inclfigure{fig:components}{components.pdf}{
    Relationship between different client-side components of the \zbf
    peer-to-peer download service. Components in gray are standard browser
    APIs.
}

\subsection{Connection manager}

The connection manager is responsible for establishing connections between
peers and facilitating peer-to-peer message transfers. It is a thin layer on
top of WebRTC that handles maintaining connections with multiple peers,
associating peers with peer IDs that are assigned by our tracker. It manages
all signaling through the signaling channel, which implements signaling through
our central server using WebSockets. At this time, this signaling server is the
same server that is running the tracker, but they are logically separate
components, so theoretically, they could be decoupled if needed for better load
balancing.

\subsection{Download manager}

The download manager is the top-level client interface for peer-to-peer
downloads. It provides an API where the client can provide a hash, and
optionally, a fallback HTTP URL, and the download manager will asynchronously
download the content associated with the hash and call a user-provided callback
function when finished. The download manager implements reliable download,
preferring peer-to-peer over HTTP when possible.

\subsubsection{Chunking}

\zbf divides assets into chunks; chunks are individually requested and
sent between peers. We use a standard chunk size of 10240 bytes in our
implementation; preliminary tests showed this chunk size to be reasonable.
When a client asks for a resource, the download manager for that client
determines the number of chunks for that resource from the specified size.
The download manager then evenly distributes requests for these chunks
among the peers with the resource. \\
Requests for chunks time out on a per-chunk basis. We track the time \zbf
last sent a request for each chunk; if this time exceeds a cutoff, we randomly
choose another peer with the resource and send the chunk request to that
peer. We periodically retrieve an updated peers list for this resource
from the tracker.

\subsubsection{Hash validation}

When fetching content peer-to-peer, it is necessary to authenticate data that
is downloaded. \zbf achieves this by using a cryptographic hash function to
authenticate data that is downloaded. Assets that are to be downloaded
peer-to-peer are identified by their hash, and once the download manager
finishes downloading the content from peers, it verifies that the hash of the
contents matches what is expected. The download manager rejects content if
there is an authentication error. In the current implementation, if this
occurs, the download manager falls back to downloading the content over HTTP
from the fallback URL.

\subsubsection{Parallelization}

\zbf divides assets into chunks primarily to enable parallelization. Because
individual peers may not have high outbound bandwidth, it is beneficial to
download in parallel from many peers. Our implementation parallelizes chunk
downloads, with pending chunk requests balanced between available peers in the
network. If certain peers are slow to respond, the download manager
automatically requests chunks from faster peers to optimize download time.
Parallel downloads help reduce download latency.

\subsubsection{Fault tolerance}

Our download manager implements reliable download semantics. The implementation
is resilient to peer failure, automatically re-requesting chunks from live
peers when necessary. In the worst case, even if all peers go down, the
implementation falls back to downloading over pure HTTP from the centralized
server as a last resort. In any case, the download manager guarantees that the
data will be downloaded and that the download will be authenticated.

\subsection{Tracker interface}

Via the tracker interface, the download manager can request a list of peers that
have a certain resource, specifying the desired resource via that resource's
unique SHA1 hash. After a client receives the full resource, the download
manager publicizes that this client has the resource via this tracker interface,
letting the central tracker know that it possesses the resource.

\subsection{Signaling channel}

The signaling channel facilitates opening the WebRTC data channel connection between
peers. The WebRTC protocol requires one peer to send a WebRTC offer and ICE candidate
to the other peer, and that other peer to send back a WebRTC answer, for a peer-to-peer
data channel to be opened. The server-side signaling channel facilitates this
information transfer between peers by  directing offers, ICE candidates, and answers
to their specified peer recipient.

\subsection{DOM injection layer}
%! Ankush

\section{Server implementation}

There are two main server-side components to \zbf. First is the tracker,
which maintains a mapping from resource hashes to the peers that have that
resource. Second is the WebRTC handshake channel, which passes information
between peers to enable them to open WebRTC data channels.

\subsection{Tracker}
The tracker is hosted on a remote server and tracks client connections.
The client-side tracker interface  opens a WebSocket connection with the remote tracker,
which registers the existence of this client. The remote tracker responds to
requests from the client for lists of peers with a given resource. When a client
publicizes that it has a resource through the tracker interface, the remote tracker updates
the list of peers with the resource to include this client, as
long as its WebSocket connection remains open. Upon WebSocket disconnection,
the tracker ``forgets'' about the client and erases all data about the client's
held resources.

\subsection{WebRTC handshake channel}

WebRTC handshake information--the offers, ICE candidates, and answers--are
sent through the client-side signaling channel through a websocket to
the server. The client-side signaling channel specifies a peer to which
this WebRTC handshake information should be sent; if the server has an open
websocket connection with that peer, the server passes that message to the peer.

\section{Evaluation}

% TODO

\section{Future work}

\subsection{Per-chunk hash validation}
As of right now, \zbf verifies the integrity of the downloaded resource once all chunks have been retrieved and joined. In normal circumstances where all chunks successfully download, this may not prove to be an issue. However, waiting until all chunks are downloaded and combined to verify the file means that if a single chunk had been corrupted or compromised, all chunks for the file would need to be downloaded, combined, and verified again.

An improvement to \zbf that would mitigate this issue would be to have the Tracker module provide checksums for each chunk. From the user's perspective, the \texttt{data-zf-hash} attribute currently used to store the hash of the file would be replaced with the hash of the JSON object storing a list of the hashes for each chunk. Upon connection to the tracker, each client would receive that list and verify its integrity, and then be able to verify and redownload each chunk independently in the event of failure..

\subsection{Hybrid downloads over HTTP and P2P}

Another improvement to \zbf would be to allow it to parallelize downloads via HTTP and P2P transfers simultaneously. By using the \texttt{Range} HTTP header attribute, \zbf could fetch the same data as contained in P2P chunks from the HTTP fallback server. By utilizing both HTTP and P2P, \zbf may be able to provide service speed that is expected in all instances to be faster than simply using HTTP alone.

\subsection{Variable request sizes}

Implementing variation in requested chunk sizes could allow \zbf to independently
determine a more optimal chunk size for each asset. For instance, we hypothesize
that different chunk sizes are optimal for differently sized assets, and potentially
for different media types. If \zbf more intelligently determined the optimal size
of chunks for each resource, overall download latency would decrease.

\subsection{Bandwidth detection and prioritization}

If \zbf could detect which peers have higher bandwidth connections, and thus
are more capable of swiftly delivering data to a client, \zbf could prioritize
requesting data from these peers. Perhaps \zbf could request larger chunks from
peers with higher bandwidth connections. \zbf could also dynamically react to
poor connections with peers during the chunk requesting process; chunk requests
that subsequently time out could be redistributed more intelligently among
peers with higher bandwidth connections.

\subsection{Support for HTML5 media}
It is possible to add functionality such that \zbf can construct \texttt{ReadableStream}s (as specified by the Streams standard) %TODO: cite https://streams.spec.whatwg.org
from the chunks of data as it receives them. Using this stream, the DOM injection layer could feasibly replace HTML5 media elements such as embedded Video and Audio. Since these types of media are particularly large in size, being able to distribute them in a decentralized, peer-to-peer manner could provide substantial load reduction for the central server.

\subsection{Automatic conversion to \zbf}
% TODO

\section{Conclusions}

% Katie

{\footnotesize \bibliographystyle{ieeetr}
\bibliography{paper}}

\end{document}